---
title:  "Testing parallel in R"
date:   "October 2015"
output: 
    html_document:
        fig_width:  15
        fig_height: 4
---

```{r libs_and_global_options, include = FALSE}
library(knitr)
opts_chunk$set(results="hold", warning=FALSE, message=FALSE, cache=TRUE) ## change to FALSE to re-eval whole doc
library(dplyr); library(magrittr); library(ggplot2)
library(parallel)
library(numbers)  # for the isPrime function
```

This note has two purposes:

- to test R's `parallel` package's performance on a basic multicore machine
- to test a wrapper function around `base::lapply` and `parallel::parLapply`

### 1. Definition of varLapply

My objective for `varLapply` is to be a no-hassle wrapper around `lapply` and `parLapply`, with a syntax similar to `lapply` and a simple boolean `use_parallel` argument that could be switched on or off algorithmically.

```{r varLapply_definition}
varLapply <- function(X, FUN,               # same syntax as base::lapply
                      use_parallel = TRUE,  # use parLapply by default
                      number_of_nodes = detectCores()-1,  # keep 1 core free by default
                      par_apply_function = "parLapply",   # allows to switch to mclapply
                      ...) {
    if (!use_parallel) {
        lapply(X, FUN, ...)
    } else {
        require(parallel)
        if (par_apply_function == "parLapply") {
            tmp_cluster <- makeForkCluster(nnodes = max(number_of_nodes,1))
            clusterExport(cl = tmp_cluster, varlist = c(), envir = parent.frame())
            output <- parLapply(cl = tmp_cluster, X = X, fun = FUN, ...)
            stopCluster(tmp_cluster)
            return(output)
        } else if (par_apply_function == "mclapply") {
            mclapply(X = X, FUN = FUN, mc.cores = number_of_nodes)
        }
    }
}
```

The purpose of `clusterExport(... envir = parent.frame())` is to import all variables/functions/objects from `varLapply`'s parent. There are probably more subtle ways of doing this, but it has worked so far.

### 2. Test 1: isPrime

```{r prime_test_definition`}
library(dplyr); library(magrittr); library(ggplot2)
library(parallel)
library(numbers)  # for the isPrime function

prime_test <- function(N,   # a vector, the integers we test for primality
                       K,   # a vector, the lengths of the [l/parL/mcl]apply loops
                       P) { # an integer, the number of times each test is repeated for averaging
    N.l <- length(N)
    K.l <- length(K)
    S.l <- N.l * K.l * P * 2
    tests <- data.frame(matrix(NA, nrow = S.l, ncol = 5))
    colnames(tests) <- c("n", # the integer tested for primality
                         "k", # the number of times n is tested
                         "p", # the number of times each test is repeated
                         "b", # if (b), uses a parallel version of lapply
                         "time") # the measured test time (elapsed in system.time)
    i <- 1
    for (n in 1:N.l) {
        for (k in 1:K.l) {
            for (p in 1:P) {
                for (b in c(FALSE, TRUE)) {
                    tests[i,] <- c(N[n], K[k], p, b, 0.0)
                    i <- i+1
    }}}}
    # each row of tests corresponds to a test unit where we run
    # isPrime(n) k times using varLapply(..., use_parallel = b);
    # output of system.time (elapsed) is stored in the time column;
    # each case is repeated P times for averaging;
    # we sample test units in random order to reduce bias potential
    for (j in sample(1:S.l)) { 
        tests[j,"time"] <- system.time(varLapply(X = rep(tests[j,"n"], tests[j,"k"]),
                                                 FUN = isPrime,
                                                 use_parallel = tests[j,"b"], 
                                                 par_apply_function = "mclapply")
                                       )[3]
        gc()
    }
    return(tests)
}
```

```{r run_test_1, cache=FALSE}
set.seed(123456)
rdm_primes <- Primes(10^2,10^9) %>% sample(size = 1000)
loop_sizes <- c(10,25,50,100,250,1000)
test_1     <- prime_test(N = rdm_primes, K = loop_sizes, P = 5)
save(test_1, file = "./test_2.Rdata")
```








---

Links that I have found useful:

- http://stackoverflow.com/questions/12019638/using-parallels-parlapply-unable-to-access-variables-within-parallel-code
- https://trinkerrstuff.wordpress.com/2012/08/19/parallelization-speed-up-functions-in-a-package/
- https://stat.ethz.ch/R-manual/R-devel/library/parallel/doc/parallel.pdf
- http://journal.r-project.org/archive/2012-2/RJournal_2012-2_Baaaath.pdf

Links that look useful and I still need to read:
- http://blog.dominodatalab.com/simple-parallelization/

---

R version:
```{r, echo=FALSE}
version[c("platform", "os", "version.string")]
```

```{r}
detectCores()
```

*Author: Alexandre Halm*  
*Edited with [RStudio](http://www.rstudio.com/products/RStudio/#Desk)*  